---
  title: "Setting up prediction environment"
author: "Orestis"
date: "28/06/2020"
output: 
  bookdown::html_document2: 
  toc : true
toc_float : true
editor_options: 
  
  chunk_output_type: console
---
  
  <style>
  p.caption {
    font-size: 0.9em;
    font-style: italic;
    color: grey;
    margin-right: 10%;
    margin-left: 10%;  
    text-align: center;
  }
</style>
  
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)

library("tidyverse")
library("lubridate")

#rm(list = ls())

source("ImportLong.r")


```

# Conceptual

Can we use a linear (or log linear) model as the basis for forecasting? 
For example, when it comes to forecasting new cases, researchers typically use variations of the SIR model. 

# Split into two periods: increase and decrease of SI

The following code, keeps the days of SI weakly increasing for every country.

```{r}
w1<-df%>%
  filter(!is.na(StringencyIndex))%>%
  group_by(Country)%>%
  mutate(day=row_number())%>%  ###Putting a counter for each row of each group
  add_tally()%>% ## creating a variable "n" that counts how many rows each group has.
  arrange(Date)%>% 
  mutate(wave1=if_else(StringencyIndex>=dplyr::lag(StringencyIndex),1,0))%>%
  mutate(fir=match(0,wave1)-1)%>%### first occurence of a 0-value - 1 to discard the 0. 
  mutate(fir=tidyr::replace_na(fir,5))%>% ### countries that never had a 0 appear as NA. We replace NAs with total rows of 1's
  mutate(wave=ifelse(day<=fir,"first","second"))%>%
  filter(wave=="first")%>%
  ungroup()


#Test
w1%>%
  group_by(Country)%>%
  ggplot(aes(x=week,y=StringencyIndex))+
  geom_line(size=1)+
  facet_wrap(~Country)

# df4<-df3%>%
#   arrange(Country,Date)%>%
#   dplyr::select(Country,n,wave1,StringencyIndex,Date,fir)


```
OK, it works. 


# Next steps:
 
1. calculate compliance coefs on first wave. 
  1.1 different counties have different lengths of wave 1. How does that affect our analysis? Duration of wave 1 brings about differences in power but it is also behavioual factor.
2. Create the coefficient table and a regression throwin in the kitchen sink. 
  2.1 Can the target set be different for every country?


