---
title: "Covid-19"
output: bookdown::html_document2
#output: html_document
editor_options: 
  chunk_output_type: console
---

```{r echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height = 6, fig.width = 12)


library(tidyr)
library(dplyr)
library(stringr)
library(lmerTest)
library(stargazer)
library(ggplot2)
library(mfx)
library(knitr)
library(kSamples)
library(FSA)
library(kableExtra)
library(xtable)
library(miceadds)
library(jtools)
library(plm)
library(clubSandwich)
library(ggrepel)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(naniar)

options("jtools-digits" = 4,scipen=999) # set displayed decimal places to 4

df<-read.csv(here::here("24042020_short.csv"))
dfl<-read.csv(here::here("22042020_long.csv"))


#df corresponds to the short version of the data, where we keep only the last observed date by Google. 
#dfl corresponds to the long version of the data, where we have day by day observations for Google and for the weather.
#The weather starts recording from 22-01-2020 while Google from mid Feb. 
#In the short version we take average temperatures and the last day of Google. 




#breaks<-df%>%
 # filter(Popular==1)%>%
  #dplyr::select(Country)





```

# Data description
I compiled the data set by collecting country-level variables from a host of sources. 

* google mobility reports: displays the change in visits to places like grocery stores and parks as of 29/03/2020 (the report was released on the 2nd of April I think but it covers up to 3 days before). 
  * The beginning of the observation period (compared to the benchmark) is Feb. 23. This seems to be consistent across all countries.
  * The baseline is the median value, for the corresponding day of the week, during the 5- week period Jan 3–Feb 6, 2020.
  * There is a new update on the 9th of April. They will keep on updating. 
* official lock-down dates for each country (various sources).
* dates of first death and 1st reported case. I have also the dates for each additional death and case-confired. 
* world bank measures for GDP per capita, unemployment and other health related indexes. 
* Rule of Law from 2018 (this is a measure that was also included in the antisocial punishment paper by Gachter et al.)
* Number of tests and deaths - though this is sth that needs to be updated ever so often

* from World Value Scales: values such as the importance of democracy and respect for ind. human rights nowadays
* average temperature and humididty for the period between 22nd of January to March 24 (thanks to World Weather Online AP)
* demographics such as median age, population, population density, urban density
* behavioural variables such as the Rule of Law (from World Bank), risk taking propensity, altruism, patience, etc. (from the global preferences survey). 


Some important dates to consider:

* 11/03/2020 : The World Health Organisation declares COVID-19 a pandemic. 
* 21-22/01/2020: First official reports conceding there are human-to-human transmissions.
* 23/02/2020: Start of observation period for mobility restrictions.
* 29/03/2020: end of observation period for 1st Google report. 
* 05/04/2020: end of observation period for 2nd Google report. 
* 22/01/2020 - 24/03/2020: period of weather observation

The idea was to explore the ability of measures such as such the rule of law, importance of democracy and respect of individual rights to correlate with a) policy decisions such as when to apply restriction measures and how severe those are and b) with individual compliance: how much was mobility restricted. 

Disclaimer: this is not an attempt to identify causal relations. This is just a report to satisfy curiocities that were raised over coffee-breaks. If we are to make a more serious scientific claims to motivate this, I guess we can claim the following two-fold contribution:
1. Assessing the usefulness of social-value indexes such as the ROL, etc. 
2. Present a case for future studies, with more ambitious identification strategies, that aim to infer causality. 





# Movement analysis

```{r mobility, fig.cap="Movement restriction across countries"}


df%>%
  filter(!is.na(Movement))%>%
 mutate(Country = forcats::fct_reorder(Country,-Movement)) %>%

  ggplot()+
  geom_point(aes(x=Country,y=Movement),colour="blue")+

 # scale_y_discrete()+
  #scale_x_discrete()+
  labs(x = 'Countries',y = "Movement")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```

I focus on the dependent variable: "Movement" for the moment. 
This variable is based on Google reports. I aggregate a single score per country by taking the average on all categories (Parks, Groceries...) except "residential". Residential measures: "Movement trends for places of residence". Unlike the other categories, residential goes up during this period (people typically move around their houses more). 

Movement probably captures two (latent for the moment) variables: a. governments' stricness of enforcing the measures b. citizens' obedience to governments' mandate. 
Although the source is reliable, it's prob. good to keep in mind that there are strong selection issues: the variable captures the behaviour of people who use smartphones and have allowed Google to track their location. 


I present below the results of two regressions. 
I label the first one: "Interesting" (table \@ref(tab:tab-model-table)) because it features two behavioural variables that correlate sigificantly with mobility: ROL and risktaking. 
The interpretation of "risktaking" is intuitive. People that are more risk-seeking restrict their movement less. 
The ROL's sign is counter-intuitive at face value. 
Countries that score higher in ROL quarantine less. Perhaps, ROL, captures other attitudes as well (such as importance of individual liberties etc.). 



```{r }


vars_intr<-c("Lock","log_GDP_pc","DaysDuration","Days_LockDown_1death","ROL","HumidityDif","Humidity",
               "TemperatureDif","Temperature","Unemployment","Pop_Density_sq_mi","risktaking","patience","ROL","polity2")

intr<-formula(paste("Movement ~", paste(vars_intr, sep = "", collapse = "+")))
intr_mob<-lm(data=df,intr)
tab_model(intr_mob, title = "(\\#tab:tab-model-table) Determinants of Movement: 'Interesting'")




```



The second regession: "Best", scores much higher in R_sq_adjusted (more than 50%) and has stronger power (more countries in the pool). The reason why the two regressions differ wrt the number of observations is because some variables are recorded for only a smaller subset of countries (such as risktaking and patience).


```{r regMov_bst, fig.cap="Determinants of Movement: Best_R" ,results='asis'}
vars_best<-c("Lock","log_GDP_pc","DaysDuration","Days_LockDown_1death","ROL","HumidityDif","Humidity",
               "TemperatureDif","Temperature","Unemployment","Death_disease","Pop_Density_sq_mi","polity2")

best<-formula(paste("Movement ~", paste(vars_best, sep = "", collapse = "+")))
best_mob<-lm(data=df,best)
tab_model(best_mob, title = "(\\#tab:tab-model-table) Determinants of Movement: 'Best'")


```



In principle, the really important explanatory variables, should persist (i.e. be significant) in both regressions. 
Unless of course, there is a pattern in the countries that drop out. In principle, this should not be the case as Briq claims to have a representative sample of the world in its measurements. Let's see, which countries are included in Google's reports but not in Briq:

```{r}

q<-df%>%
  filter(!is.na(Movement),is.na(risktaking))%>%
  dplyr::select(Country)

print(q$Country)

#kable(q$Country, booktabs=T,title = "(\\#tab:MoveNBriq) Google:Yes - Briq:No") 

```



Death_disease is a peculiar case for two reasons.^[I found the data for this variable in World Bank's data set and it tracks: "Cause of death, by communicable diseases and maternal, prenatal and nutrition conditions (% of total)".] 
First, the sign suggests that countries with higher %'s of communicable diseases (i.e. one that is spread from one person to another through a variety of ways that include: contact with blood and bodily fluids; breathing in an airborne virus; or by being bitten by an insect), tend to restrict their movement less during this pandemic. 
Second, including this variable crowds out the effect of GDP_pc. 
Perhaps this variable correlates too strongly with GDP_pc so the effect that is picking up is misleading. 
Figure \@ref(fig:GDPFlu) suggests that this is more likely the case. 


```{r GDPFlu, fig.cap="GDP_pc across Death by flu"}


df%>%
  
 #mutate(Country = forcats::fct_reorder(Country,-Movement)) %>%

  ggplot(aes(x=log(GDP_pc),y=Death_disease),colour="blue")+
  #geom_vline(xintercept=-12.7)+
  #geom_hline(yintercept=-48.6)+
  #geom_point(aes(size=before_Google,colour=Lock),alpha=0.4)+
  #guides(colour="legend")+
  stat_smooth()+


 # scale_y_discrete()+
  scale_x_discrete()+
  labs(x = 'GDP_pc',y = "Death by communicable disease (% total)")+
  theme_bw()+
  geom_text_repel(aes(x=log(GDP_pc),y=Death_disease,label=Country))+
 #geom_text(x=-7.5,y=5,label="High Death Rate - High Movement", vjust = "inward", hjust = "inward")+
  #geom_text(x=-16,y=-90,label="Low Death Rate - Low Movement", vjust = "inward", hjust = "inward")+

  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  #facet_wrap(~Lock)


```

Goals:

* Find a proxy variable for how much a country likes to touch, hug and kiss each other. 
* What crowds out ROL? 







# Death and total cases. 

```{r LogDeathpc, fig.cap="Death by country"}


df%>%
  filter(Death_pc>0,Population>1000000)%>%
 mutate(Country = forcats::fct_reorder(Country,log(Death_pc))) %>%

  ggplot()+
  geom_point(aes(x=Country,y=log(Death_pc)),colour="blue")+

 # scale_y_discrete()+
  #scale_x_discrete()+
  labs(x = 'Countries',y = "Log_Death_pc")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



```


What does restriction of movement do to total cases of death? 
Hard to answer this question. 

Figure \@ref(fig:MobDeath) plots the relation between Total Deaths per capita and Movement. 
The relation is endogenous. In principle, one would expect that lower movelemt -> less deaths. 
But of course, it may be that the countries that took the lock-down more seriously are the ones that were afflicted the most (higher death toll). 
I estimate the variable before_Google, that counts how many deaths were officially confirmed before Google's observation period began. 
I estimate this variable as: 

$$
log(\frac{(Deaths(before23/02/2020)+1)}{Population})
$$

I add +1 in the numerator, because there are many countries with 0 reported deaths before google, and logging leads to -Inf. 
Not sure if this is the right approach, maybe I should let them drop out. 




```{r MobDeath, fig.cap="Movement restriction and Death"}


df%>%
  mutate(Google_1=Google+1)%>%
 mutate(Google_1pc=Google_1/Population)%>%
  mutate(before_Google=log(Google_1pc))%>%
  
  filter(!is.na(Movement))%>%
# mutate(Country = forcats::fct_reorder(Country,-Movement)) %>%

  ggplot(aes(x=Log_Death_pc,y=Movement),colour="blue")+
  geom_vline(xintercept=-12.7)+
  geom_hline(yintercept=-48.6)+
  geom_point(aes(size=before_Google,colour=Lock),alpha=0.4)+
  #guides(colour="legend")+
  stat_smooth()+


 # scale_y_discrete()+
  scale_x_discrete()+
  labs(x = 'Log_Death_pc',y = "Movement")+
  theme_bw()+
  geom_text_repel(aes(x=Log_Death_pc,y=Movement,label=Country))+
 geom_text(x=-7.5,y=5,label="High Death Rate - High Movement", vjust = "inward", hjust = "inward")+
  geom_text(x=-16,y=-90,label="Low Death Rate - Low Movement", vjust = "inward", hjust = "inward")+

  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  #facet_wrap(~Lock)


```






It's interesting to look at those 4 quadrants. 
First, notice that quadrants 1 and 3 (low deaths) are typically comprised of countries that are less likely to be accurately reporting deaths (limited interest and/or medical supplies). 
Perhaps we can control for that with a proxy for health-services (e.g. hospital beds per capita). 

Is there a story about the second quadrant? 
Nordic countries seem to dominate this quadrant of high-movement & high death rate. 
Is this is a story of climate or can we fit an explanation with a political science twist (e.g. value of individual liberties)? 

Moreover, how do we control for the direction of causality? 
It's unlikely that reducing movement increases total deaths. 
More likely, the countries that were stricter at reducing mobility were the ones that were mostly affected by Covid.
I create the before_Google variable that tracks down the number of deaths attributed to covid-19 before the observation period started. 
We would expect that this control will reduce the effect of mobility on total deaths. 



```{r regSmpl, fig.cap="Determinants of Death_toll: Simplistic" ,results='asis'}



smpl<-lm(data=df,Log_Death_pc~Movement)


tab_model(smpl)




```


```{r regCntrl, fig.cap="Determinants of Death_toll: add before_Google control" ,results='asis'}

r1<-df%>%
  mutate(Google_1=Google+1)%>%
 mutate(Google_1pc=Google_1/Population)%>%
  mutate(before_Google=log(Google_1pc))%>%
  
  filter(!is.na(Movement))


ctrl<-lm(data=r1,Log_Death_pc~Movement+before_Google)
tab_model(ctrl)

```

The coefficient of Movement is reduced but not nearly as much as I would have thought...
Maybe the amount of deaths before Google's observation period is not a good predictor. 

```{r MobDeathBefore, fig.cap="Movement restriction and Death before Google"}


df%>%
  mutate(Google_1=Google+1)%>%
 mutate(Google_1pc=Google_1/Population)%>%
  mutate(before_Google=log(Google_1pc))%>%
  
  filter(!is.na(Movement))%>%
# mutate(Country = forcats::fct_reorder(Country,-Movement)) %>%

  ggplot(aes(x=before_Google,y=Movement),colour="blue")+
  #geom_vline(xintercept=-12.7)+
  #geom_hline(yintercept=-48.6)+
  geom_point(aes(colour=Lock),alpha=0.4)+
  #guides(colour="legend")+
  stat_smooth()+


 # scale_y_discrete()+
  scale_x_discrete()+
  labs(x = 'before_Google',y = "Movement")+
  theme_bw()+
  geom_text_repel(aes(x=before_Google,y=Movement,label=Country))+
 geom_text(x=-7.5,y=5,label="High Death Rate - High Movement", vjust = "inward", hjust = "inward")+
  geom_text(x=-16,y=-90,label="Low Death Rate - Low Movement", vjust = "inward", hjust = "inward")+

  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  #facet_wrap(~Lock)


```





# Flattening the curve

Just started playing around with dynamic trends.


```{r trend, fig.cap="Death curve"}

dfl%>%
  mutate(Date=as.Date(Date))%>%
  mutate(label = if_else(Date == max(Date,na.rm=T), as.character(Country), NA_character_)) %>%

  filter(TotalDeaths>1000,Date>as.Date("01/03/2020",format="%d/%m/%y"))%>%
  group_by(Country)%>%
  ggplot(aes(x = Date, y = total_deaths,group=Country, color=Country)) + 
  geom_line() +
  #scale_colour_brewer(palette = 'Set1')+
  #guides(colour="legend")
  theme_minimal()+
  scale_color_discrete(guide = FALSE)+
  geom_label_repel(aes(label = label),
                  nudge_x = 1,
                  na.rm = TRUE)




```




# Action points

* Find the country (region) - level individualism/ collectivism (see Hofstede, Gelfland, etc.) & integrate.
* Distance from lockdown in days.
* glmnet selects the variables that together offer the best predictive power. In the interesting model we see what I would think is somewhat counter-intuitive variables (like patience:in, lokc:out). Let's run the models suggested by glmnet and compare. 
  * Lock is actually out for a technical reason which we should prob. fix. 
* Finish Literature review on individual and collectivism
* Integrate the behavioural data and strictness of lock-down?
  * Data structure: let's make sure it works. 
  * careful in integrating the strictness of lock-down. Multiconearity (do they take into account the Movement reduction for eg?). 
  * Orestis: behavioural data and strictness of lock-down, collectivism indexes
  * Lio: Polity4, UN-Population...emdat.be
  
* What is the Death_Disease about? 
  





