---
title: "Results of ARDL estiations with covariates"
author: "Lio, Orestis"
date: "03/09/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("sjPlot")

source("ARDL_fun.R")

```

## Intro

This report provides an overview of the results obtained from the ARDL estimations that account for climatic variables as developed on 26th of August (aka 'last Wednesday').

Brief summary of the model: up to 7 lags (1 week) for both dependent and independent variable, dependent variable is movement, independent variables are stringency index, precipitation and temperature. 

Note 1: Only units with at least 50 observations were retained, as otherwise there would be computational errors (I suppose degrees of freedom exhausted with 7*4 variables as well as constant, etc.)

Note 2: We take first differences of all series, which is because they are very obviously non-stationary.

Note 3: Computational errors arose from constant precipitation series in dry regions. We added a random small error term to the series as a quick fix. A cleaner solution would be to confirm if there is variation for each country before estimating, and to remove the variable if there isn't any.

## Data

To simplify / accellerate things, I saved the results of the estimations in code/ARDL_cov_03092020.Rds.

The data comes as a list of 5 items:

```{r load data, echo = TRUE}

ARDL_cov <- readRDS("ARDL_cov_03092020.Rds")
str(ARDL_cov,1)

```
The first 3 elements are lists of matrices: e.g., bgtest has a 7*7 matrix for each country, reporting the p-value of the breush-godfrey test for all combinations of lag-lengths p and q.

Item 4 stores the entire model for each city, that includes all coefficients and their SEs.

Item 5 is a dataframe with one entry per city. It reports the key statistics for each city and the model that was chosen for it:

```{r, echo = TRUE}

head(ARDL_cov$summary)

```


## Key Results

First, let's look at what kinds of models our criteria like:

```{r lang length, echo=FALSE}

ARDL_cov$summary %>%
  summarise(average.p = mean(p.best),
            average.q = mean(q.best))

```


The choice seems to be with models that include few lags of the independent variables, but a lot of autoregressive ones (typically maxing out). Indeed, increasing p is much more 'costly' in terms if degrees of freedom, as it applies to 3 variables. This could in principle be made more flexible, but at a high cost (not readily implemented in the functions we make use of).

Now let's look at the parameter of interest, the long run coefficient:

```{r, echo=TRUE}

summary(ARDL_cov$summary$LongRunCoefficient)
```

Nothing super-weird here, a 1 unit increase in stringency is on average associated with a 0.44% decrease in mobility as measured by google.

There are 411 coefficients, so perhaps they are best summarised in the form of a histogram/density plot.



```{r lrcoeff, echo=FALSE, fig.height=16, fig.width=8}
df.plot<- ARDL_cov$summary %>%
  mutate(ObsUnit = forcats::fct_reorder(ObsUnit,-LongRunCoefficient))

ggplot(df.plot,
       aes(x = LongRunCoefficient)) + 
  geom_histogram(aes(y = ..density..), fill = "grey") +
  geom_density(color = "red") +
  theme_bw() 

```

Only few coefficients are in the weird positive spectrum, to be precise `r sum(ARDL_cov$summary$LongRunCoefficient>0)`.

## Correlates of LR coefficients

Let's turn to potential correlates of those long-run coefficients.

```{r correlates prep, echo=FALSE, warning=FALSE,message=FALSE}
setwd("..")
source("ImportLong.r")

df2<-df%>%
  filter(!is.na(StringencyIndex),!is.na(Movement))
  
  

library("plyr")
#Break up w1 by country, then fit specified model to each piec and return a list
models<-dlply(df2,'Country',function(df)
  lm(Movement~StringencyIndex,data=df))

#apply coef to each model and return a data frame
q<-ldply(models,coef)
names(q)<-c("Country","Intercept","Compliance")


#breaks<-df%>%
 # filter(Popular==1)%>%
  #dplyr::select(Country)

df_avg<-df%>%
 filter(!is.na(Movement),!is.na(StringencyIndex))%>%
  group_by(Country)%>%
  summarise_at(vars(-c(total_deaths,total_cases)),mean,na.rn=TRUE)%>%
  ungroup()

df_other<-df%>%
   filter(!is.na(Movement),!is.na(StringencyIndex))%>%
  group_by(Country)%>%
  dplyr::summarise(total_deaths=last(total_deaths),total_cases=last(total_cases),LogDeath=(log(last(total_deaths)+1)),LogCases=log(last(total_cases)+1),MaxSI=max(StringencyIndex,na.rm=TRUE))%>%
  ungroup()

df_short<-merge(df_avg,df_other,by="Country",all=T)  
df_short<-merge(df_short,q,by="Country",all=T) 
df_short<-merge(df_short,output.ardl$summary[,c("ObsUnit","LongRunCoefficient")],by.x="Country",
                by.y = "ObsUnit",all=T)

vars_basic<-c("log_GDP_pc","Unemployment","Pop_Density_sq_mi","Age_65up","Death_disease")
vars_basic_GDP<-c("Unemployment","Pop_Density_sq_mi","Age_65up","Death_disease")
vars_briq<-c("risktaking","patience","negrecip","trust","posrecip","altruism")
vars_hof<-c("COL","PDI","MAS","UAI")
```




```{r regressions, echo=FALSE}

basic<-formula(paste("LongRunCoefficient ~", paste(vars_basic, sep = "", collapse = "+")))
basic_mob<-lm(data=df_short,basic)

briq<-formula(paste("LongRunCoefficient ~", paste(c(vars_basic,vars_briq,"ROL"), sep = "", collapse = "+")))
briq_mob<-lm(data=df_short,briq)

hof<-formula(paste("LongRunCoefficient ~", paste(c(vars_basic,vars_briq,"ROL",vars_hof), sep = "", collapse = "+")))
hof_mob<-lm(data=df_short,hof)

tab_model(basic_mob,briq_mob,hof_mob)

```

Somewhat different to earlier estimates; not sure how intuitive, needs closer consideration.

A different angle:


```{r boxplot, echo=FALSE, message=FALSE, warning=FALSE}

library("mosaic")

df.plot <- df_short %>%
  mutate(Regime = derivedFactor(
    "Autocracy" = polity2 <= -6,
    "Anocracy" = (polity2 > -6 & polity2 < 6),
    "Democracy" = polity2 >= 6)
    )

ggplot(df.plot, aes(y=LongRunCoefficient,colour = Regime)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```

Immediate observation is that there is a huge diversity among democracies; maybe this can be fine tuned by removing outliers or so.


```{r boxplot2, echo=FALSE}

library(countrycode)

df.plot$Continent <- countrycode(df.plot$Country,origin = "country.name", destination = "continent")

ggplot(df.plot, aes(y=LongRunCoefficient,colour = Continent)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```

Perhaps this plots would be informative anyway, also for other metrics (such as average movement, for instance).
