---
title: "Covid-19: Scoping the field"
author: "Orestis, Lio"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2: 
    toc : true
    toc_float : true
editor_options: 
  chunk_output_type: console
---

```{r echo = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height = 6, fig.width = 12)

library(tidyverse)
library(stringr)
library(lmerTest)
library(stargazer)
library(mfx)
library(knitr)
library(kSamples)
library(FSA)
library(kableExtra)
library(xtable)
library(miceadds)
library(jtools)
library(plm)
library(clubSandwich)
library(ggrepel)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(naniar)
library(ggpubr)
library(lubridate)
library(ggmap)
library(tufte)

options("jtools-digits" = 4,scipen=999) # set displayed decimal places to 4


rm(list = ls())



gwmp<-readRDS("gwmp.rds")

city_short<-gwmp%>%
  filter(!is.na(Movement),!is.na(risktaking),!is.na(Temp_C))%>%
  group_by(Country,City)%>%
  summarise(
     across(where(is.numeric), mean), 
     across(where(is.factor), first),
     across(where(is.character), first),
     n = n())

country_short<-city_short%>%
  group_by(Country)%>%
   summarise(
     across(where(is.numeric), mean), 
     across(where(is.factor), first),
     across(where(is.character), first),
     n = n())
  


```

# Introduction

> The ultimate measure of a man is not where he stands in moments of comfort and convenience, but where he stands at times of challenge and controversy.
`r tufte::quote_footer('---  Martin Luther King')`


Our goal in this project is to construct a new behavioural measure of 'compliance'. 
Unlike most other behavioural approaches that elicit such measures with experimental procedures, our approach is to derive this measure from data capturing real-life responses. 
We treat Covid-19 as an exogenous shock that allows us to track real-world behaviour, namely, responsiveness to Policy (but also, to each country's respective urgency).
The high-stakes (life or death) of the situation and the attention it has received world-wide make this index all the more meaningful.


# Key variables


## Stringency Index

This derived from a quantitative integration of governmental responses to the the pandemic. 

### Oxford

### Coronanet

## Movement

Describe Google's mobility reports. 

# Some tables and graphs

Notice how I round the digits now from the table.

```{r}
DT::datatable(
  city_short%>%
    dplyr::select(Country,City,Temp_C,risktaking, patience, Movement, population,n)
   
)%>%  
  DT::formatRound(columns=c('Temp_C','risktaking','patience','Movement'),digits=3)
  

```


# Missing variables: Weather

We need to find a way to cite. 

Going more granular: city level analysis. 

Increasing the level of our data-granularity is a good thing. 
Even if we cluster errors at the country-level, we still get a boost on power. 
Moreover, it does not matter if we do not get city-level data for all variables of interest. 
For example, @alfaro2020social, use city-level mobility data but country-level death and total cases data. 
We probably should do the same. 


On the map:

```{r}
qmplot(lon, lat, data = city_short, maptype = "toner-lite", color = I("red"))

```
